Debiased Calibration for Generalized Two-Phase Sampling

### 2024-07-29 final/reports/prelim_paper.tex
Updated July 29, 2024

\begin{theorem}[Variance Estimation]
  The variance of $\hat Y_{NNE}$ is

@@

where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
%\textcolor{red}{The simulation setup is somewhat strange. The first order
%inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a
%function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
%In this case, your second-phase sampling is essentially equivalent to the simple
%random sampling. }
We consider a finite population of size $N = 10,000$ with the Phase 1 
sampling being a simple random sample (SRS) of size $n_1 = 1000$. The Phase 2
sample is a Poisson sample with an expected sample size of

@@

where $\hat T = (\hat{\bf X}^{(0)}, \sum_{i \in U} g(d_i) q_i)^T$ and 
$\bf z_i = (\bf x_i^T, g(d_i)q_i)^T$. A more general approach considers the case
in which multiple samples can observe $Y$. In this case, assume that we have
samples $A_1, \dots A_{M_1}$ for which we observe elements of $\bf X$ and $Y$
and samples $B_1, \dots, B_{M_2}$ for which we only observed elements of $\bf
X$. We denote $\bf X^{(A_m)}$ and $\bf X^{(B_m)}$ as the columns of $\bf X$
observed in samples $A_m$ and $B_m$ respectively. Then the final
estimator\footnote{This is sort of like stratified sampling with unknown strata.}
is

$$ \hat Y = M_1^{-1} \sum_{m = 1}^{M_1} \sum_{i \in A_m} \hat w_{im} y_i $$

@@

where $\hat T = (\hat{\bf X}^{(A_m)}, \sum_{i \in U} g(d_{im}) q_i)^T$ and $d_{im}$
are the design weights for $A_m$.

@@

>>>>

\begin{theorem}[Variance Estimation]\label{thm:nnve}
  The variance of $\hat Y_{NNE}$ is

@@

where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
%\textcolor{red}{The simulation setup is somewhat strange. The first order
%inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a
%function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
%In this case, your second-phase sampling is essentially equivalent to the simple
%random sampling. }
We consider a finite population of size $N = 10,000$ with the Phase 1 
sampling being a simple random sample (SRS) of size $n_1 = 1000$. The Phase 2
sample is a Poisson sample with an expected sample size of about 300.

@@

where $\hat T = (\hat{\bf X}^{(0)}, \sum_{i \in U} g(d_i) q_i)^T$ and 
$\bf z_i = ((\bf x_i^{(0)})^T, g(d_i)q_i)^T$. Then the estimator is $\hat Y_{MS}
= \sum_{i \in A_0} w_{i} y_i$.

A more general approach considers
the case in which multiple samples can observe $Y$. In this case, assume that we
have samples $A_1, \dots A_{M_1}$ for which we observe elements of $\bf X$ and
$Y$ and samples $B_1, \dots, B_{M_2}$ for which we only observed elements of
$\bf X$. We denote $\bf X^{(A_m)}$ and $\bf X^{(B_m)}$ as the columns of $\bf X$
observed in samples $A_m$ and $B_m$ respectively. Then the final
estimator\footnote{This is sort of like stratified sampling with unknown
strata.} is

$$ \hat Y_{MSE} = M_1^{-1} \sum_{m = 1}^{M_1} \sum_{i \in A_m} \hat w_{im} y_i $$

@@

where $\hat T = (\hat{\bf X}^{(A_m)}, \sum_{i \in U} g(d_{im}) q_i)^T$, $\bf
z_i^{T} = ((\bf x_i^{(A_m)})^T, g(d_{im})q_i)^T$ and $d_{im}$
are the design weights for $A_m$.

@@

We prove the theory for the simpler case in which we have samples $A_0, A_1,
\dots, A_{M}$.

\begin{theorem}[Design Consistency]
  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$.
  Under regularity conditions,

  $$\hat Y_{MS} = \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) + O_p(N / n_0)$$

  $$\hat Y_\ell(\bm \lambda^*, \bm \phi^*) = \hat Y_{DCE} + \left(
    \hat T - \sum_{i \in A_0} \hat w_{0i}(\bm \lambda^*) \bf z_i q_i
  \right)\bm \phi^*$$

  $$\bm \phi^* = \left[\sum_{i \in U} \frac{\pi_{0i} \bf z_i \bf z_i^T
  q_i}{g'(d_{0i})}\right]^{-1} \sum_{i \in U} \frac{\pi_{0i} \bf z_i
  y_i}{g'(d_{0i})}.$$
\end{theorem}

\begin{proof}
  This proof follows from the proof of Theorem~\ref{thm:dc2}. The only
  difference is that we have $\hat X_{GLS}$ instead of $\hat X_c$, but the final
  result still holds.
\end{proof}

Since this result holds for each simple case, and our general case is the
average of each of the simple cases that we observe, the result holds for
Equation~\eqref{eq:dc32} as we assume each $A_m$ is independent of each other.

\begin{theorem}[Variance Estimation]
  Under regularity conditions,

  $$
  \begin{aligned}
    V(\hat Y_{MS}) 
    &= M_1^{-2} \left(\sum_{m = 1}^{M_1} 
    \Big\{(\bm \phi_1^{(m)*})^T\Var(\hat{\bf X}_{GLS}^{(m)})
      (\bm \phi_1^{(m)*})^T \right.\\
    &+ \left.
      \sum_{i \in U} \sum_{j \in U} \frac{\Delta_{mij}}{\pi_{mi}\pi_{mj}}
      (y_i - \bf z_i^{(m)}\bm \phi^{(m)*}q_i)
      (y_j - \bf z_j^{(m)}\bm \phi^{(m)*}q_j) \Big\} \right).
  \end{aligned}
  $$

  We can estimate the variance with

  $$
  \begin{aligned}
    \hat V(\hat Y_{MS}) 
    &= M_1^{-2} \left(\sum_{m = 1}^{M_1} 
      \Big\{(\hat{\bm \phi_1}^{(m)})^T\hat \Var(\hat{\bf X}_{GLS}^{(m)})
        (\hat{\bm \phi_1}^{(m)})^T \right.\\
    &+ \left.
      \sum_{i \in A_m} \sum_{j \in A_m} \frac{\Delta_{mij}}{\pi_{mij}\pi_{mi}\pi_{mj}}
      (y_i - \bf z_i^{(m)}\hat{\bm \phi}^{(m)}q_i)
      (y_j - \bf z_j^{(m)}\hat{\bm \phi}^{(m)}q_j) \Big\} \right).
  \end{aligned}
  $$

\end{theorem}

\begin{proof}
  This result follows the same argument as Theorem~\eqref{thm:nnve}. The result
  holds because each sample $A_m$ is independent from the other $A_{m'}$. Since,
  the score equation for the regression estimator asserts that $\sum_{i \in A_m}
  \bf z_i (y_i - \bf z_i \hat{\bm \phi}^{(m)}) = \bf 0$, the covariance terms
  should be approximately zero.
\end{proof}
